# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r, echo=TRUE, eval=TRUE}
unzip('activity.zip')
accel <- read.csv('activity.csv', header=TRUE, sep = ',')
str(accel)
for(i in 1:length(accel) ) {
  str(accel[i])
  }
```

We observe that the **date** variable is of the type **Factor**, we therefore recode it as a **date** variable.

```{r echo=TRUE, eval=TRUE}
accel$date <- as.Date( accel$date, format = '%Y-%m-%d' )
str(accel$date)
```

## What is mean total number of steps taken per day?
We start by loading the packages for data manipulation and plotting.

```{r echo=TRUE, eval=TRUE }
library(plyr)
library(ggplot2)
```

We now compute the data needed for the plot.

```{r echo=TRUE, eval=TRUE }
tot.steps <- ddply(accel, .(date),
                  summarize,
                  total.steps = sum(steps)
                  )
head(tot.steps)
```

We can now plot the data.

```{r histogram, echo=TRUE, eval=TRUE }
tot.st.plot <- ggplot(tot.steps,
                      aes(x = total.steps)
                      )
tot.st.plot <- tot.st.plot + geom_histogram()
print(tot.st.plot)
```

We now report the summary statistics, to see the mean and median.

```{r echo=TRUE, eval=TRUE }
summary(tot.steps$total.steps)
summary(tot.steps$total.steps)['Mean']
summary(tot.steps$total.steps)['Median']
```



## What is the average daily activity pattern?
Compute the data needed for the plot.

```{r echo=TRUE, eval=TRUE }
av.steps <- ddply(accel, .(interval),
                  summarize,
                  average.steps = mean(steps, na.rm=TRUE
                                       )
                  )
head(av.steps)
```

Plot the computed data.

```{r plot, echo=TRUE, eval=TRUE }
av.st.plot <- ggplot(av.steps,
                     aes(interval, average.steps)
                     )
av.st.plot <- av.st.plot + geom_line()
print(av.st.plot)
```

Find the maximum.

```{r echo=TRUE, eval=TRUE }
summary(av.steps$average.steps)
summary(av.steps$average.steps)['Max.']
```



## Imputing missing values

First we count the rows that have NAs and the ones that do not.

```{r echo=TRUE, eval=TRUE}
table( complete.cases(accel) )
```

Create a new data frame and impute missing values using interval averages.

```{r echo=TRUE, eval=TRUE}
imp.accel <- accel
imp.accel$steps <- ifelse( is.na(imp.accel$steps),
                     av.steps$average.steps[match(imp.accel$interval, av.steps$interval) ],
                     imp.accel$steps )
table( complete.cases(accel) )
table( complete.cases(imp.accel) )
```

```{r echo=TRUE, eval=TRUE }
tot.imp.steps <- ddply(imp.accel, .(date),
                  summarize,
                  total.steps = sum(steps)
                  )
head(tot.imp.steps)
```

We can now plot the data.

```{r imputed.histogram, echo=TRUE, eval=TRUE }
tot.imp.st.plot <- ggplot(tot.imp.steps,
                      aes(x = total.steps)
                      )
tot.imp.st.plot <- tot.imp.st.plot + geom_histogram()
print(tot.imp.st.plot)
```

We now report the summary statistics, to see the mean and median.

```{r echo=TRUE, eval=TRUE }
summary(tot.imp.steps$total.steps)
summary(tot.imp.steps$total.steps)['Mean']
summary(tot.imp.steps$total.steps)['Median']
```


## Are there differences in activity patterns between weekdays and weekends?


